WSClean: an implementation of a fast, generic wide-field imager

### 2014-03-01 articles/2014/Article-WSClean/wsclean.tex
nwlayer 6->12
Slight fixes

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{lrrr}%
\hline\hline
& \textbf{\textsc{wsclean}} & \textbf{\textsc{wsclean}}    & \textbf{\textsc{casa}} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (6 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & --- & 1.90 mJy\\
Computational time & 8.5 min & & 19.3 min\\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (128 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.39\% & & 2.08\% \\
RMS in residual image & 0.94 mJy & --- & 0.94 mJy \\
Computational time & 10.3 min & & 19.6 min \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 10\degree (195 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.75\% & 1.40\% & 2.41 \%\\
RMS in residual image & 0.90 mJy & 1.03 mJy & 1.07 mJy \\
Computational time & 15.3 min & 6.6 min & 178.2 min \\
\hline\hline
\end{tabular}\end{center}\end{table}

@@

\begin{table*}
\caption{Predicted computational costs for configurations listed in Table~\ref{tbl:computational-cost-configurations}, based on multi-frequency synthesis with five major iterations observing for 1~h at ZA=20\degree with 1 polarization. Predictions are for $w$-projection with \textsc{casa}; $w$-stacking with \textsc{wsclean}; $w$-stacked snapshots with \textsc{wsclean} using optimal snapshot duration; and a hybrid between $w$-projection and $w$-stacking using \textsc{casa} with optimal $\Delta w$. All have approximately equal accuracy.} \label{tbl:computational-cost-predictions}
\begin{tabular}{l|rrrr|rr}
               & \multicolumn{4}{c|}{Predicted computing cost on test computer} & min & best \\
 Configuration & $w$-projection & $w$-stacking & $w$-snapshot & hybrid & computing & FLOPS/float \\
 \hline
 GLEAM    &  65h 25m & 8h 04m &  8h 03m & 14h 10m   & 1.1 TFLOPS & $6.9 \times 10^2$    \\
 EMU      &    5.2 d &  2.9 d &   2.9 d & 4.4 d     & 9.7 TFLOPS & $2.1 \times 10^4$      \\
 MSSS low &  2h 16m  & 0h 57m &  0h 57m & 2h 10m    & 130 GFLOPS & $3.5 \times 10^3$   \\
 MSSS high&  7h 16m  & 3h 52m &  3h 52m & 5h 44m    & 530 GFLOPS & $3.4 \times 10^3$   \\
 LOFAR NL &  50.2 d  &  7.3 d &   7.2 d & 12.9 d    &  24 TFLOPS & $4.5 \times 10^3$ \\
 AARTFAAC &  2.5 d   &  1.2 d &   1.2 d &   2.5 d   & 4.1 TFLOPS & $6.8 \times 10^2$   \\
 VLSS     &  0h 09m  & 0h 01m &  0h 01m &  0h 08m   & 2.2 GFLOPS & $12.1 \times 10^3$     \\
 MeerKAT  &   10.8 d &  6.2 d &   6.2 d &  10.8 d   &  21 TFLOPS & $6.8 \times 10^2$ \\
% SKA1 AA inner&392 d &  228 d &   228 d &   392 d & & $6.8 \times 10^2$   \\
 SKA1 AA core&1581 d &  911 d &   911 d &  1570.8   & 3.0 PFLOPS & $6.8 \times 10^2$ \\
 SKA1 AA full& 643 yr& 48.8 yr & 48.8 yr& 84.1 yr   &  59 PFLOPS & $6.8 \times 10^2$ \\
\end{tabular}
\end{table*}

@@

\subsection{Estimated computing cost for other telescopes} \label{sec:application-to-nonmwa}
We use the derived functions to estimate the computational cost of the algorithms for configurations of several telescopes. We estimate the cost of imaging an hour of ZA=20\degree data for the following surveys or array configurations: The Galactic and Extragalactic MWA survey (GLEAM); the Evolutionary Map of the Universe (EMU) ASKAP survey \citep{emu-norris-2011}; the low and high bands of the LOFAR Multi-frequency Snapshot Sky Survey (MSSS)\footnote{See~\href{https://www.astron.nl/radio-observatory/lofar-msss/lofar-msss}{www.astron.nl/radio-observatory/lofar-msss/lofar-msss}}; all Dutch LOFAR stations \citep{lofar-2013}; the Amsterdam--ASTRON Radio Transients Facility and Analysis Centre (AARTFAAC) project\footnote{See \href{http://www.aartfaac.org}{www.aartfaac.org}}; The VLA Low-Frequency Sky Survey (VLSS, \citealt{vlss-2007}); The Meer Karoo Array Telescope (MeerKAT); and the low-frequency Phase~1 aperture arrays of the Square-Kilometre Array (SKA), with the full core (3~km) and the core+arms configurations \citep{ska-phase1-2013}. The configurations are summarized in Table~\ref{tbl:computational-cost-configurations}. In multi-beam configurations such as the SKA and EMU configurations, we assume each beam is imaged separately, i.e., the FOV of a single beam is used and the computing cost is multiplied with the number of beams. The selected wavelengths 
are approximately the central wavelength available for each instrument. The image size is set to two times the FOV width divided by the resolution, as described by Eq.~\ref{eq:nyquist-resolution}. The total required computing power is calculated by multiplying the estimated computing time on our test machine with the performance of the machine (138 GFLOPS).

@@

\textsc{wsclean} is currently not able to perform multi-scale clean. Results of applying \textsc{casa}'s multi-scale clean on MWA data show that especially the Galactic plane is significantly better deconvolved using multi-scale clean, but it is very expensive. Imaging a one-minute observation takes 30 hours of computational time without $w$-projection. We plan to implement some form of multi-scale cleaning in \textsc{wsclean}. It is likely that additional optimizations need to be made to be able to do this with acceptable performance.

>>>>

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{lrrr}%
\hline\hline
& \textbf{\textsc{wsclean}} & \textbf{\textsc{wsclean}}    & \textbf{\textsc{casa}} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (12 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & --- & 1.90 mJy\\
Computational time & 8.5 min & & 19.3 min\\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (128 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.39\% & & 2.08\% \\
RMS in residual image & 0.94 mJy & --- & 0.94 mJy \\
Computational time & 10.3 min & & 19.6 min \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 10\degree (195 $w$-layers/planes)}} \\
\hline
Source flux standard error & 1.75\% & 1.40\% & 2.41 \%\\
RMS in residual image & 0.90 mJy & 1.03 mJy & 1.07 mJy \\
Computational time & 15.3 min & 6.6 min & 178.2 min \\
\hline\hline
\end{tabular}\end{center}\end{table}

@@

\begin{table*}
\caption{Predicted computational costs for configurations listed in Table~\ref{tbl:computational-cost-configurations}, based on multi-frequency synthesis with five major iterations observing for 1~h at ZA=20\degree with 1 polarization. Predictions are for $w$-projection with \textsc{casa}; $w$-stacking with \textsc{wsclean}; $w$-stacked snapshots with \textsc{wsclean} using optimal snapshot duration; and a hybrid between $w$-projection and $w$-stacking using \textsc{casa} with optimal $\Delta w$. All have approximately equal accuracy.} \label{tbl:computational-cost-predictions}
\begin{tabular}{l|rrrr|rr}
               & \multicolumn{4}{c|}{Predicted computing cost on test computer} & min & best \\
 Configuration & $w$-projection & $w$-stacking & $w$-snapshot & hybrid & computing & FLOPS/float \\
 \hline
 GLEAM    &  65h 25m & 8h 04m &  8h 03m & 14h 10m   & 1.1 TFLOPS & $6.8 \times 10^2$    \\
 EMU      &    5.2 d &  2.9 d &   2.9 d & 4.4 d     & 9.7 TFLOPS & $2.1 \times 10^4$      \\
 MSSS low &  2h 16m  & 0h 57m &  0h 57m & 2h 10m    & 130 GFLOPS & $3.4 \times 10^3$   \\
 MSSS high&  7h 16m  & 3h 52m &  3h 52m & 5h 44m    & 530 GFLOPS & $3.4 \times 10^3$   \\
 LOFAR NL &  50.2 d  &  7.3 d &   7.2 d & 12.9 d    &  24 TFLOPS & $7.2 \times 10^2$ \\
 AARTFAAC &  2.5 d   &  1.2 d &   1.2 d &   2.5 d   & 4.1 TFLOPS & $6.8 \times 10^2$   \\
 VLSS     &  0h 09m  & 0h 01m &  0h 01m &  0h 08m   & 2.2 GFLOPS & $1.1 \times 10^3$     \\
 MeerKAT  &   10.8 d &  6.2 d &   6.2 d &  10.8 d   &  21 TFLOPS & $6.8 \times 10^2$ \\
% SKA1 AA inner&392 d &  228 d &   228 d &   392 d & & $6.8 \times 10^2$   \\
 SKA1 AA core&1581 d &  911 d &   911 d &  1570.8   & 3.0 PFLOPS & $6.8 \times 10^2$ \\
 SKA1 AA full& 643 yr& 48.8 yr & 48.8 yr& 84.1 yr   &  59 PFLOPS & $6.8 \times 10^2$ \\
\end{tabular}
\end{table*}

@@

\subsection{Estimated computing cost for other telescopes} \label{sec:application-to-nonmwa}
We use the derived functions to estimate the computational cost of the algorithms for configurations of several telescopes. We estimate the cost of imaging an hour of ZA=20\degree data for the following surveys or array configurations: The Galactic and Extragalactic MWA survey (GLEAM); the Evolutionary Map of the Universe (EMU) ASKAP survey \citep{emu-norris-2011}; the low and high bands of the LOFAR Multi-frequency Snapshot Sky Survey (MSSS)\footnote{See~\href{https://www.astron.nl/radio-observatory/lofar-msss/lofar-msss}{www.astron.nl/radio-observatory/lofar-msss/lofar-msss}}; all Dutch LOFAR stations \citep{lofar-2013}; the Amsterdam--ASTRON Radio Transients Facility and Analysis Centre (AARTFAAC) project\footnote{See \href{http://www.aartfaac.org}{www.aartfaac.org}}; The VLA Low-Frequency Sky Survey (VLSS, \citealt{vlss-2007}); The Meer Karoo Array Telescope (MeerKAT); and the low-frequency Phase~1 aperture arrays of the Square-Kilometre Array (SKA), with the full core (3~km) and the core+arms configurations \citep{ska-phase1-2013}. The configurations are summarized in Table~\ref{tbl:computational-cost-configurations}. In multi-beam configurations such as the EMU and LOFAR configurations, we assume each beam is imaged separately, i.e., the FOV of a single beam is used and the computing cost is multiplied with the number of beams. The selected wavelengths 
are approximately the central wavelength available for each instrument. The image size is set to two times the FOV width divided by the resolution, as described by Eq.~\ref{eq:nyquist-resolution}. The total required computing power is calculated by multiplying the estimated computing time on our test machine with the performance of the machine (138 GFLOPS).

@@

\textsc{wsclean} is currently not able to perform multi-scale clean. Results of applying \textsc{casa}'s multi-scale clean \citep{multiscale-clean-cornwell-2008} on MWA data show that especially the Galactic plane is significantly better deconvolved using multi-scale clean, but it is very expensive. Imaging a one-minute observation takes 30 hours of computational time without $w$-projection. We plan to implement some form of multi-scale cleaning in \textsc{wsclean}. It is likely that additional optimizations need to be made to be able to do this with acceptable performance.
